---
title: "Statistical Learning- Supervised project"
author: "Mina Beric"
date: "2024-05-10"
output: html_document
---

IMPORT REQUIRED LIBRARIES

```{r setup, include=FALSE}

library(tidyverse)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(patchwork)
library(corrplot)
library(knitr)
library(superml)
library(caret)
library(magrittr)
library(nnet)
library(MLmetrics)
library(DAAG)
library(party)
library(rpart)
library(rpart.plot)
library(mlbench)
library(pROC)
library(tree)
library(car)
library(glmnet)
library(randomForest)
library(mgcv)
library(adabag)
library(gtsummary)
```

```{r cars}
data1=read.csv("C:/Users/beric/Downloads/archive/ObesityDataSet_raw_and_data_sinthetic.csv", sep=",")
data1
str(data1)
glimpse(data1)
summary(data1)
```

In this dataset we have 2111 observations and 17 variables,

Data cleaning and pre processing

Finding null values

```{r pressure, echo=FALSE}
anyNA(data1)
```

no null values

Rename the attributes to become more intuitive
```{r}
data1<-rename(data1, High_cal_food_freq=FAVC,
                     Veg_in_meals=FCVC,
                     N_main_meal_daily=NCP,
                     Snacks=CAEC,
                     Water_daily_l= CH2O,
                     Calory_monitor=SCC,
                     Physical_activity_days=FAF,
                     Tech_use_hours=TUE,
                     Alcohol_consump=CALC,
                     Transportation=MTRANS,
                     Obesity_level=NObeyesdad)
head(data1)
```

Make some changes in the dataset for visualization reasons
```{r}
data1 <- data1 %>%
  mutate(Transportation = ifelse(Transportation == "Public_Transportation", "Public_transp", Transportation))

```
As this project compares the effectiveness of different classification techniques, I will transform the output variable to a binary output.
For this problem:
Change Obesity_level into binary for the purpose of the analysis:
```{r}
data1 <- data1 %>%
  mutate(Obesity_level = case_when(Obesity_level %in% c("Normal_Weight", "Overweight_Level_I", "Overweight_Level_II", "Insufficient_Weight") ~ "ObeseNo",
                                   Obesity_level %in% c("Obesity_Type_I", "Obesity_Type_II", "Obesity_Type_III") ~ "ObeseYes",
                                   TRUE ~ Obesity_level  # Keep unchanged if not matching any of the above
  ))

data1<- data1 %>%
  mutate(Obesity_level=factor(ifelse(Obesity_level=="ObeseYes",1,0))) 

str(data1$Obesity_level)

```

See if there are some Duplicates, if yes remove them 
```{r}
sum(duplicated(data1)) 

#Remove the duplicates
Data1_unique<-distinct(data1)
Data1_unique
tail(Data1_unique)
```
Some records of the continuous variables exhibit an unexpected number of decimal places, the potential causes could be Data entry Errors or measurment inaccuracy. 
Let's handle this problem by transforming them into integers:
```{r}
Data1_unique$Age<- as.integer(Data1_unique$Age)
Data1_unique$Veg_in_meals<- as.integer(Data1_unique$Veg_in_meals)
Data1_unique$N_main_meal_daily<-as.integer(Data1_unique$N_main_meal_daily)
Data1_unique$Water_daily_l<-as.integer(Data1_unique$Water_daily_l)
Data1_unique$Physical_activity_days<-as.integer(Data1_unique$Physical_activity_days)
Data1_unique$Tech_use_hours<-as.integer(Data1_unique$Tech_use_hours)

tail(Data1_unique)
```
Let's handle the categorical variables
 
transform categorical variables into factorial 
```{r}
Data1_unique<-Data1_unique %>%
  mutate(Gender=factor(Gender),
         Alcohol_consump=factor(Data1_unique$Alcohol_consump, levels=c("no","Sometimes","Frequently","Always")),
         High_cal_food_freq=factor(High_cal_food_freq),
         Calory_monitor=factor(Calory_monitor),
         SMOKE=factor(SMOKE),
         family_history_with_overweight=factor(family_history_with_overweight),
         Snacks=factor(Snacks, levels=c("no","Sometimes","Frequently","Always")),
         Transportation=factor(Transportation))

```
let's verify the changes
```{r}
glimpse(Data1_unique)
```
UNIVARIATE ANALYSIS of the independent variables:
```{r}
summary(Data1_unique)
attach(Data1_unique)
```
Starting from the categorical variables: 
```{r}
#1. Gender

ggplot(Data1_unique, aes(x = Gender, fill = Gender)) +
  geom_bar() + 
  labs(title = "Gender Distribution", x = "Gender", y = "Count") +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)+
  scale_fill_manual(values = c("pink", "lightblue")) + 
  theme(legend.position = "none")

#2. Alcohol consumption

ggplot(Data1_unique, aes(x = Alcohol_consump,fill=Alcohol_consump)) +
  geom_bar() +
  labs(title = "Distribution of Alcohol consumption habits",
       x = "Alcohol_consump",
       y = "Frequency") +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_brewer(palette="Set3")

ggplot(Data1_unique, aes(x = "", fill = Alcohol_consump)) +
  geom_bar(width = 1) +
  coord_polar(theta = "y",start=0) +  
  labs(title = "Distribution of Alcohol Consumption Habits",
       fill = "Alcohol Consumption",
       x = NULL,  
       y = NULL) +  
  geom_text(aes(label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_stack(vjust = 0.5), 
            size = 3,
            angle = c(-90, -90, 0, 0)) +
  scale_fill_brewer(palette = "Set3") 


#3. High Calory food frequency consumption

ggplot(Data1_unique, aes(x = High_cal_food_freq, fill = High_cal_food_freq)) +
  geom_bar() + 
  labs(title = "High calory consumption frequency", x = "High calory consump", y = "Count") +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 4)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values = c("darkgreen", "darkred")) + 
  theme(legend.position = "none")


#4. Calory monitor: 

ggplot(Data1_unique, aes(x = Calory_monitor, fill = Calory_monitor)) +
  geom_bar() + 
  labs(title = "Calory monitoring frequency", x = "Calory monitoring", y = "Count") +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 4)+
  scale_fill_manual(values = c("darkred", "darkgreen")) + 
  theme(legend.position = "none")


#5.Smoke: 

ggplot(Data1_unique, aes(x = SMOKE, fill = SMOKE)) +
  geom_bar() + 
  labs(title = "Smoking distribution", x = "Smoking", y = "Count") +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 4)+
  scale_fill_manual(values = c("darkgreen", "darkred")) + 
  theme(legend.position = "none")

#6. Family history of being overweight: 

ggplot(Data1_unique, aes(x = family_history_with_overweight, fill = family_history_with_overweight)) +
  geom_bar() + 
  labs(title = "Family history of being overweight frequency", x = "Family history overweight", y = "Count") +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 4)+
  scale_fill_manual(values = c("darkgreen", "darkred")) + 
  theme(legend.position = "none")


#7. Snacks

ggplot(Data1_unique, aes(x = Snacks, fill = Snacks)) +
  geom_bar() + 
  labs(title = "Snacking between meals", x = "Snacking", y = "Count") +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 4)+
  scale_fill_brewer(palette="Set3")+
  theme(legend.position = "none")


ggplot(Data1_unique, aes(x = "", fill = Snacks)) +
  geom_bar( width = 1) +
  coord_polar("y", start = 0, direction = -1) + 
  labs(title="Distribution of Snacking habits",
       fill="Snacking habits",
       x=NULL,
       y=NULL)+
  geom_text(aes(label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_stack(vjust = 0.5), 
            size = 3,
            angle = c(-90,0,-330 , 0)) +
  scale_fill_brewer(palette = "Set3") 
  
   

#8. Transportation 

ggplot(Data1_unique, aes(x = Transportation, fill = Transportation)) +
  geom_bar() + 
  labs(title = "Transportation frequency", x = "Transportation", y = "Count") +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 4)+
  scale_fill_brewer(palette="Set3") + 
  theme(legend.position = "none")


ggplot(Data1_unique, aes(x = "", fill = Transportation)) +
  geom_bar(width = 1) +
  coord_polar("y", start = 0, direction = -1) + 
  labs(title="Distribution of Transportation od choice",
       fill="Transportation",
       x=NULL,
       y=NULL)+
  geom_text(aes(label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_stack(vjust = 0.5), 
            size = 3,
            angle = c(0,0,0,0,0)) +
  scale_fill_brewer(palette = "Set3") 

```
Relations between categorical variables
```{r}
#Alcohol consumption vs Gender

  Data1_unique %>% 
  count(Alcohol_consump, Gender) %>%  
  ggplot(mapping = aes(x = Alcohol_consump, y = Gender)) +
  geom_tile(mapping = aes(fill = n))+
    scale_fill_gradient(low="lightblue", high="darkred")

  Data1_unique %>%
    count(Alcohol_consump,Gender) 
  
  
#Alcohol consumpt vs Smoke
  
  Data1_unique %>% 
    count(Alcohol_consump, SMOKE) %>%  
    ggplot(mapping = aes(x = Alcohol_consump, y = SMOKE)) +
    geom_tile(mapping = aes(fill = n))+
    
    scale_fill_gradient(low = "lightblue", high = "darkred") 
    
  Data1_unique %>%
    count(Alcohol_consump,SMOKE) 

#Gender vs Smoke
  
  Data1_unique %>% 
    count(Gender, SMOKE) %>%  
    ggplot(mapping = aes(x = Gender, y = SMOKE)) +
    geom_tile(mapping = aes(fill = n))+
    scale_fill_gradient(low = "lightblue", high = "darkred") 
  
  Data1_unique %>%
    count(Gender,SMOKE) #prevalebce of non smoker for both genders

  
#High calory food frequency vs Calory monitor
  
  Data1_unique %>% 
    count(High_cal_food_freq, Calory_monitor) %>%  
    ggplot(mapping = aes(x = High_cal_food_freq, y = Calory_monitor)) +
    geom_tile(mapping = aes(fill = n))+
    scale_fill_gradient(low = "lightblue", high = "darkred") 
  
  Data1_unique %>%
    count(High_cal_food_freq,Calory_monitor)
  
#High calory food frequency vs Family history overweight
  
  Data1_unique %>% 
    count(High_cal_food_freq, family_history_with_overweight) %>%  
    ggplot(mapping = aes(x = High_cal_food_freq, y = family_history_with_overweight)) +
    geom_tile(mapping = aes(fill = n))+
    scale_fill_gradient(low = "lightblue", high = "darkred") 
  
  Data1_unique %>%
    count(High_cal_food_freq,family_history_with_overweight) 
  
  #prevalence of people that have a high calory diet also has family history overweight
  
#family_history_with_overweight vs Gender
  
  Data1_unique %>% 
    count(family_history_with_overweight, Gender) %>%  
    ggplot(mapping = aes(x =family_history_with_overweight, y = Gender)) +
    geom_tile(mapping = aes(fill = n))+
    scale_fill_gradient(low = "lightblue", high = "darkred") 
  
  Data1_unique %>%
    count(family_history_with_overweight, Gender) 
  
```
Continuous variables: 
```{r}
#9. Age

plot1 <- ggplot(Data1_unique, aes(x = Age)) +
  geom_histogram(aes(y = ..density..), bins = 30, alpha = 0.5, fill = "green") +
  geom_density(aes(y = ..density..), linetype = "solid", color = "darkgreen") +
  labs(title = "Density of Age", x = "Age", y = "Density") +
  scale_y_continuous(name = "") +
  theme_classic()

plot2 <- ggplot(Data1_unique, aes(x = Age)) +
  geom_boxplot(aes(y = Age), notch = TRUE, color = "black") +
  labs(title = "Boxplot of Age", x = "Age", y = "Values") +
  theme_classic()

grid.arrange(plot1, plot2, nrow = 1)

#Age by Gender groups
Data_copy<-Data1_unique
Data_copy$age_group <- cut(Data_copy$Age, c(0,20,24,35,70), labels = c("<21","21-24","25-35",">35") )


ggplot(Data_copy,aes(x=age_group, fill=Gender))+
  geom_bar(col="black")+
  facet_wrap(.~Gender)+
  stat_count(aes(y=..count.., label=..count..), vjust=-0.5,geom="text", col="black", size=3.5)+
  labs(x="Age Group", y = "Count", title="Age distribution", fill= "Sex")+
  theme_minimal()+
  theme(plot.title=element_text(face="bold",  hjust=0.5))+
  scale_fill_manual(values=c("pink","lightblue"))

#Age vs weight by gender
ggplot(Data1_unique, aes(x = Age, y = Weight, color = Gender)) +
  geom_point() +  
  labs(title = "Weight vs. Age by Gender",  
       x = "Age",  
       y = "Weight")  

 
#11. Veg_in_meals


plot5<- ggplot(Data1_unique, aes(x = Veg_in_meals))+
  geom_histogram(aes(y = ..density..),bins=3, alpha = 0.5, fill = "orange") +
  geom_density(aes(y = ..density..), linetype = "solid", color = "red") +
  labs(title = "Density of Vegetables in meals", x = "Veg_in_meals", y = "Density") +
  scale_y_continuous(name = "") +
  theme_classic()

plot6<-ggplot(Data1_unique, aes(x = Veg_in_meals)) +
  geom_boxplot(aes(y = Veg_in_meals), notch = TRUE, color = "black") +
  labs(title = "Boxplot of Vegetables in meals", x = "Veg_in_meals", y = "Values") +
  theme_classic()

grid.arrange(plot5, plot6, nrow = 1)


#12. N_main_meals_daily

plot7<- ggplot(Data1_unique, aes(x = N_main_meal_daily)) +
  geom_histogram(aes(y = ..density..),bins=3, alpha = 1.5, fill = "pink") +
  geom_density(aes(y = ..density..), linetype = "solid", color = "darkviolet") +
  labs(title = "Density of the number of main meals per day", x = "N_main_meals_daily", y = "Density") +
  scale_y_continuous(name = "") +
  theme_classic()

plot8<-ggplot(Data1_unique, aes(x = N_main_meal_daily)) +
  geom_boxplot(aes(y = N_main_meal_daily), notch = TRUE, color = "black") +
  labs(title = "Boxplot of num of main meals per day", x = "N_main_meals_daily", y = "Values") +
  theme_classic()

grid.arrange(plot7, plot8, nrow = 1)

#13. Water_daily_l

plot9<- ggplot(Data1_unique, aes(x = Water_daily_l)) +
  geom_histogram(aes(y = ..density..),bins=3, alpha = 0.6, fill = "blue") +
  geom_density(aes(y = ..density..), linetype = "solid", color = "darkblue") +
  labs(title = "Density of Water intake in liters", x = "Water_daily_l", y = "Density") +
  scale_y_continuous(name = "") +
  theme_classic()

plot10<-ggplot(Data1_unique, aes(x = Water_daily_l)) +
  geom_boxplot(aes(y = Water_daily_l), notch = TRUE, color = "black") +
  labs(title = "Boxplot of Water intake ", x = "Water_daily_l", y = "Values") +
  theme_classic()

grid.arrange(plot9, plot10, nrow = 1)

#14. Physical_activity_days

plot11<-ggplot(Data1_unique, aes(x = Physical_activity_days, fill = Gender)) +
  geom_histogram(bins = 4, alpha = 0.7,color="black") +  # Adjust transparency for overlapping bars
  labs(title = "Distribution of Physical Activity Days by Gender", x = "Physical Activity Days", y = "Count") +
  theme_minimal() +
  scale_fill_manual(values = c("blue", "pink"))


plot12<-ggplot(Data1_unique, aes(x = Physical_activity_days)) +
  geom_boxplot(aes(y = Physical_activity_days), notch = TRUE, color = "black",fill="lightblue") +
  labs(title = "Boxplot of Physical activity habits ", x = "Physical_activity_days", y = "Values") +
  theme_classic()


grid.arrange(plot11, plot12, nrow = 1)

#14. Tech_use_hours

plot13<- ggplot(Data1_unique, aes(x = Tech_use_hours, fill=Age_Category)) +
  geom_histogram(bins=3, alpha = 0.5, color = "black") +
  geom_density(aes(y = ..density..), linetype = "solid", color = "#ffcc00") +
  labs(title = "Density of use of technology by Age_category ", x = "Tech_use_hours", y = "Density") +
  scale_y_continuous(name = "") +
  theme_minimal()


plot14<-ggplot(Data1_unique, aes(x = Tech_use_hours)) +
  geom_boxplot(aes(y = Tech_use_hours), notch = TRUE, color = "black",fill="lightgreen") +
  labs(title = "Boxplot of Technology use", x = "Tech_use_hours", y = "Values") +
  theme_classic()



#Weight vs water intake by gender
ggplot(Data1_unique, aes(x = Weight, y = Water_daily_l, color = Gender)) +
geom_point() +  # Add points to the plot
  labs(title = "Weight vs. Water intake by Gender",  
       x = "Weight", 
       y = "Water intake") 

```
```{r}
#Correlation with the target variable
data_numeric <- select_if(Data1_unique,is.numeric) 
str(data_numeric)


data_all <- cbind(data_numeric, Data1_unique$Obesity_level)
Datacopy<- Data1_unique

Datacopy$Obesity_level_numeric <- as.numeric(as.character(Datacopy$Obesity_level)) 
correlation_with_obesity <- cor(Datacopy[, c("Age", "Height", "Weight", "Veg_in_meals", "N_main_meal_daily", 
                                         "Water_daily_l", "Physical_activity_days", "Tech_use_hours", 
                                         "Obesity_level_numeric")])

corrplot(correlation_with_obesity,type = "upper", method = "number", tl.col = "black",tl.srt = 45, tl.cex = 0.8, color = colorRampPalette(c("blue", "white", "red"))(20))

```
Strong correlation between Weight and Height, this causes collinearity, remove Weight from the dataset 
```{r}
Data1_unique<-select(Data1_unique,-Weight)
str(Data1_unique)
```
boxplots and histograms of numerical variables:
```{r}
Data1_unique %>% select(-Gender,-Alcohol_consump,-High_cal_food_freq,-Calory_monitor,-SMOKE,-family_history_with_overweight,-Snacks,-Transportation,-Obesity_level) %>% gather() %>% 
  ggplot(aes(value)) + 
  geom_boxplot(fill="skyblue", color="black", outlier.colour = "blue") + 
  facet_wrap(~key, scales = 'free')+
  theme_minimal()

Data1_unique %>% select(-Gender,-Alcohol_consump,-High_cal_food_freq,-Calory_monitor,-SMOKE,-family_history_with_overweight,-Snacks,-Transportation,-Obesity_level) %>% gather() %>% 
  ggplot(aes(value)) +
  geom_histogram(fill="skyblue", color="black") + 
  facet_wrap(~key, scales = 'free')+
  theme_minimal()
```
Here is more clear, it seems there are possible outliers in the variables Age and Physical_activity_days.
I further inspect to see the nature and the entity of outliers, using IQR criterion.
```{r}
Q1 = quantile(Age,0.25)
Q3 = quantile(Age,0.75)
IQR = Q3 - Q1
lower_bound<- Q1-1.5*IQR
upper_bound<- Q3+1.5*IQR
outliers<-Data1_unique[Age<lower_bound | Age > upper_bound, ]
nrow(outliers) #147 identified

#for the Physical_activity_variable
Q1p = quantile(Physical_activity_days,0.25)
Q3p = quantile(Physical_activity_days,0.75)
IQRp=Q3p-Q1p
lower_bounds<- Q1p-1.5*IQRp
upper_bounds<- Q3p+1.5*IQRp
outliersphy<- Data1_unique[Physical_activity_days<lower_bounds | Physical_activity_days > upper_bounds, ]
nrow(outliersphy) #75 identified

#for the Height variable
Q1h = quantile(Height,0.25)
Q3h = quantile(Height,0.75)
IQRh=Q3h-Q1h
lower_bounds<- Q1h-1.5*IQRh
upper_bounds<- Q3h+1.5*IQRh
outliersheight<- Data1_unique[Height<lower_bounds | Height > upper_bounds, ]
nrow(outliersheight) #Just 1
```
Let's study the dependent variable
```{r}

summary(Data1_unique$Obesity_level)

#The distribution 

ggplot(Data1_unique, aes(x = Obesity_level,fill=Obesity_level)) +
  geom_bar() +
  labs(title = "Distribution of Obesity Levels",
       x = "Obesity Level",
       y = "Frequency") +
  geom_text(aes(y = ..count.. -200, 
                label = paste0(round(prop.table(..count..),4) * 100, '%')), 
            stat = 'count', 
            position = position_dodge(.1), 
            size = 3)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_brewer(palette="Set3")


ggplot(Data1_unique, aes(x = "", fill = Obesity_level)) +
  geom_bar(width = 1) +
  coord_polar("y", start = 0, direction = -1) + 
  labs(title="Distribution of Obesity levels",
       fill="Obesity_level",
       x=NULL,
       y=NULL)+
  geom_text(aes(label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_stack(vjust = 0.5), 
            size = 3,
            ) +
  scale_fill_brewer(palette = "Set3") 

```
```{r}
#Obesity_level vs categorical variables

categorical_cols<-c('Gender',"Alcohol_consump","High_calory_food_freq","Calory_monitor","SMOKE","family_history_with_overweight","Snacks","Transportation")

O1C<-ggplot(Data1_unique, aes(x = Gender, fill = Obesity_level)) +
  geom_histogram(stat = "count", bins = 6, alpha = 0.7) +
  labs(title = "Obesity by Gender", x = "Gender", y = "Density") +
  geom_text(aes(label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_stack(vjust = 0.5), 
            size = 3,
  ) +
  scale_fill_brewer(label=c("ObeseYes","ObeseNo"),palette="Set3") +
  theme_minimal()


O2C<-ggplot(Data_copy, aes(x = as.factor(age_group), fill = Obesity_level)) +
  geom_histogram(stat = "count", bins = 3, alpha = 0.7) +
  labs(title = "Obesity by Age Category", x = "age_group", y = "Density") +
  geom_text(aes(label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_stack(vjust = 0.5), 
            size = 3,
  ) +
  scale_fill_brewer(label=c("ObeseNo","ObeseYes"),palette="Set3") +
  theme_minimal()

O3C<-ggplot(Data1_unique, aes(x = Alcohol_consump, fill = Obesity_level)) +
  geom_histogram(stat = "count", bins = 3, alpha = 0.7) +
  labs(title = "Obesity by Alcohol consumption habits", x = "Alcohol_consumpt", y = "Density") +
  geom_text(aes(label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_stack(vjust = 0.5), 
            size = 3,
  ) +
  scale_fill_brewer(label=c("ObeseNo","ObeseYes"),palette="Set3") +
  theme_minimal()

O4C<-ggplot(Data1_unique, aes(x = SMOKE, fill = Obesity_level)) +
  geom_histogram(stat = "count", bins = 3, alpha = 0.7) +
  labs(title = "Obesity by Smoking habits", x = "SMOKE", y = "Density") +
  geom_text(aes(label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_stack(vjust = 0.5), 
            size = 3,
  ) +
  scale_fill_brewer(label=c("ObeseNo","ObeseYes"),palette="Set3") +
  theme_minimal()

O5C<-ggplot(Data1_unique, aes(x = Transportation, fill = Obesity_level)) +
  geom_histogram(stat = "count",alpha = 0.7) +
  labs(title = "Obesity by Transportation", x = "Transport", y = "Density") +
  geom_text(aes(label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_stack(vjust = 0.5), 
            size = 3,
  ) +
  scale_fill_brewer(label=c("ObeseNo","ObeseYes"),palette="Set3") +
  theme_minimal()

O6C<-ggplot(Data1_unique, aes(x = High_cal_food_freq, fill = Obesity_level)) +
  geom_histogram(stat = "count", alpha = 0.7) +
  labs(title = "Obesity  by High calory food habit", x = "High_cal_food_freq", y = "Density") +
  geom_text(aes(label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_stack(vjust = 0.5), 
            size = 3,
  ) +
  scale_fill_brewer(label=c("ObeseNo","ObeseYes"),palette="Set3") +
  theme_minimal()

O7C<-ggplot(Data1_unique, aes(x = family_history_with_overweight, fill = Obesity_level)) +
  geom_histogram(stat = "count",alpha = 0.7) +
  labs(title = "Obesity  by fam history overweight", x = "family_history_with_overweight", y = "Density") +
  geom_text(aes(label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_stack(vjust = 0.5), 
            size = 3,
  ) +
  scale_fill_brewer(label=c("ObeseNo","ObeseYes"),palette="Set3") +
  theme_minimal()

O8C<-ggplot(Data1_unique, aes(x =Snacks, fill = Obesity_level)) +
  geom_histogram(stat = "count",alpha = 0.7) +
  labs(title = "Obesity by Snacking habits", x = "Snack", y = "Density") +
  geom_text(aes(label = paste0(round(prop.table(..count..), 4) * 100, '%')), 
            stat = 'count', 
            position = position_stack(vjust = 0.5), 
            size = 3,
  ) +
  scale_fill_brewer(label=c("ObeseNo","ObeseYes"),palette="Set3") +
  theme_minimal()

grid.arrange(O1C,O2C,O3C,O4C, nrow = 2,ncol=2)
grid.arrange (O5C,O6C,O7C,O8C,nrow=2,ncol=2)


#Obesity_level vs numerical variables


numerical_vars <- c("Age", "Height","Weight" ,"Veg_in_meals", "N_main_meals_daily", 
                    "Water_daily_l", "Physical_activity_days", "Tech_use_hours")




O1N<-ggplot(Data1_unique, aes(x = Obesity_level, y = Age, fill = factor(Obesity_level))) +
  geom_violin(alpha = 0.5) + 
  labs(title = "Obesity  by Age", 
       x = "Obesity", y = "Age") + 
  theme_minimal()

O2N<-ggplot(Data1_unique, aes(x = Obesity_level, y = Height, fill = factor(Obesity_level))) +
  geom_violin(alpha = 0.5) + 
  labs(title = "Obesity  by Height", 
       x = "Obesity_level", y = "Height") +  
  theme_minimal()

#Only for EDA
#O2Nw<-ggplot(Data1_unique, aes(x = Obesity_level, y =Weight, fill = factor(Obesity_level))) + geom_violin(alpha = 0.5) + labs(title = "Obesity  by Weight",  x = "Obesity_level", y = "Weight") +  theme_minimal()

O3N<-ggplot(Data1_unique, aes(x = Obesity_level, y = Veg_in_meals, fill = factor(Obesity_level))) +
  geom_violin(alpha = 0.5) +  
  stat_summary(fun.y=median, geom="point", shape=23, size=2, fill="red")+
  labs(title = "Obesity  by Vegetables in meals", 
       x = "Obesity", y = "Vegetables in meals") + 
  theme_minimal()

O4N<-ggplot(Data1_unique, aes(x = Obesity_level, y =N_main_meal_daily, fill = factor(Obesity_level))) +
  geom_violin(alpha = 0.5) + 
  stat_summary(fun.y=median, geom="point", shape=23, size=2, fill="red")+
  labs(title = "Obesity  by Number of meals daily", 
       x = "Obesity", y = "N_main_meals_daily") +  
  theme_minimal()


O5N<-ggplot(Data1_unique, aes(x = Obesity_level, y = Water_daily_l, fill = factor(Obesity_level))) +
  geom_violin(alpha = 0.5) + 
  stat_summary(fun.y=median, geom="point", shape=23, size=2, fill="red")+
  labs(title = "Obesity  by Water_daily_l", 
       x = "Obesity", y = "Water_daily_l") +  
  theme_minimal()

O6N<-ggplot(Data1_unique, aes(x = Obesity_level, y =Physical_activity_days, fill = factor(Obesity_level))) +
  geom_violin(alpha = 0.5) +  
  stat_summary(fun.y=median, geom="point", shape=23, size=2, fill="red")+
  labs(title = "Obesity  by Physical activity", 
       x = "Obesity", y = "Physical_activity_days") +  
  theme_minimal()

O7N<-ggplot(Data1_unique, aes(x = Obesity_level, y = Tech_use_hours, fill = factor(Obesity_level))) +
  geom_violin(alpha = 0.5) +  
  labs(title = "Obesity by Technology use", 
       x = "Obesity", y = "Tech_use_hours") +  
  theme_minimal()



grid.arrange(O1N,O2N,O7N, nrow=2,ncol=2)
grid.arrange(O4N,O5N,O6N,O3N, nrow=2,ncol=2)
```
Scaling
```{r}
num_cols <- c('Age', 'Height', 'Veg_in_meals', 'N_main_meal_daily', 'Water_daily_l', 'Physical_activity_days', 'Tech_use_hours')
cat_cols <- c('Gender', 'Alcohol_consump', 'High_cal_food_freq', 'Calory_monitor', 'SMOKE', 'family_history_with_overweight', 'Snacks', 'Transportation','Obesity_level')

glimpse(Data1_unique)

cat_var<- Data1_unique[, sapply(Data1_unique, is.factor)]
glimpse(cat_var)

# Normalize numerical columns
last_col <- ncol(Data1_unique)  
num_cols_except_last <- setdiff(names(Data1_unique), names(Data1_unique)[last_col])

normalized_num <- Data1_unique[-ncol(Data1_unique)] %>%
  select(all_of(num_cols)) %>%
  scale() %>%
  as.data.frame()
normalized_num


# Combine normalized numerical and encoded categorical columns
normalized_data <- cbind(normalized_num,cat_var )

# Print the head of the preprocessed data
glimpse(normalized_data)
```
MODELS:
First, 
Divide the dataset in training and test
```{r}
set.seed(123)
sample<- sample(c(TRUE,FALSE),nrow(normalized_data),replace=TRUE,prob=c(0.7,0.3))

dtrain<-normalized_data[sample,]
dtest<-normalized_data[!sample,]


glimpse(dtrain)
tail(dtrain)
glimpse(dtest)
```
LOGISTIC REGRESSION 

Train logistic regression model
```{r}
log_reg <- glm(Obesity_level ~Age+Gender+Height+Alcohol_consump+High_cal_food_freq+Veg_in_meals+N_main_meal_daily+Calory_monitor+SMOKE+Water_daily_l+family_history_with_overweight+Physical_activity_days+Tech_use_hours+Snacks+Transportation, data = dtrain, family = "binomial") 
summary(log_reg)
exp(coef(log_reg))
results <- tbl_regression(log_reg, exponentiate=F)
results
# Predict on test set

lr_prob=predict(log_reg,dtest,type='response')
lr_pred=ifelse(lr_prob>0.5,1,0)
lr_cm=confusionMatrix(as.factor(lr_pred),as.factor(dtest$Obesity_level),positive='1')
print(lr_cm)
lr_cm$byClass

#test ROC
test_roc = roc(dtest$Obesity_level ~ lr_prob, plot = TRUE, print.auc = TRUE)

```
let's search if there is a better model


Due to the complexity of interpretation of the logistic model, I wanted to try some
shrinkage methods, in hope that it gives the possibility of excluding less relevant variables.
especially because of th large number of categorical variables

LASSO


```{r}
#Tuning best lambda

y_train=dtrain$Obesity_level
dtrain_x <- dtrain %>%
  dplyr::select(-c(Obesity_level))

x_train=model.matrix( ~ ., dtrain_x[,])
cv.lasso=cv.glmnet(x_train,y_train,alpha=1,family="binomial")
plot(cv.lasso)
print(cv.lasso$lambda.min)

#Best lambda

lasso=glmnet(x_train,y_train,alpha=1,family="binomial",lambda=cv.lasso$lambda.min)
coef(lasso)

#Accuracy

dtest_x <- dtest %>%
  dplyr::select(-c(Obesity_level))
x_test=model.matrix( ~ .,dtest_x[,])
lasso_prob=predict(lasso,x_test,type='response')
lasso_pred=ifelse(lasso_prob>0.5,1,0)
lasso_cm=confusionMatrix(as.factor(lasso_pred),as.factor(dtest$Obesity_level),positive='1')
print(lasso_cm)
```
#The accuracy is slightly better, and I've got a more simpler model (-5 variables that are not significant for predicting the level of obesity)


GAM (GENERALIZED ADDITIVE MODEL)
```{r}
obe_gam <- gam(Obesity_level ~ Gender + s(Age) + s(Height) + family_history_with_overweight + High_cal_food_freq+ Veg_in_meals + N_main_meal_daily + Snacks + SMOKE + Water_daily_l + Calory_monitor + Physical_activity_days + Tech_use_hours + Alcohol_consump + Transportation, data = dtrain, method="REML", family="binomial")
summary(obe_gam)

# predict

gam_prob=predict(obe_gam,dtest,type='response')
gam_pred=ifelse(gam_prob>0.5,1,0)

#confusion matrix
gam_cm=confusionMatrix(as.factor(gam_pred),as.factor(dtest$Obesity_level),positive='1')
print(gam_cm)
gam_cm$byClass

#ROC CURVE
#test ROC
test_roc = roc(dtest$Obesity_level ~ gam_prob, plot = TRUE, print.auc = TRUE)
```
#accuracy of approximately 82.36% on the test data. In other words, for 82.36% of the data points, the model correctly predicted the obesity level (either obese or not obese).

DECISION TREE MODEL: 
```{r}
large_tree=rpart(Obesity_level~., data=dtrain, method="class", model=TRUE)
cp=which.min(large_tree$cptable[, "xerror"]) %>% large_tree$cptable[., "CP"]
print(large_tree$cptable)
tree=prune(large_tree,cp=cp)
rpart.plot(tree)

tree_pred=predict(tree, dtest, type='class')
tree_cm=confusionMatrix(as.factor(tree_pred),as.factor(dtest$Obesity_level),positive='1')
tree_cm$byClass #to see metrics, Sensitivity, Specificity, Balanced Accuracy
print(tree_cm)
```
RANDOM FOREST:
```{r}
set.seed(777)
dtrain$Obesity_level=as.factor(dtrain$Obesity_level)
possible_mtrys=seq(1:8)
for (v_mtry in possible_mtrys) {
  class_rf=randomForest(Obesity_level~.,data=dtrain,ntree=500,mtry=v_mtry)
  print(class_rf)
}

#number of trees

set.seed(777)
class_rf_bestmtry=randomForest(Obesity_level~.,data=dtrain,ntree=500,mtry=4)
plot(class_rf_bestmtry)
legend(x="topright",box.col="black",bg="white",box.lwd=2,title="err.rate",legend=c("OOB","0","1"),fill=c("black","red","green"))


#Model and accuracy

class_rf=randomForest(Obesity_level~.,data=dtrain,ntree=300,mtry=4,importance=TRUE)
rf_predicts=predict(class_rf,dtest,type='class')
rf_cm=confusionMatrix(as.factor(rf_predicts),as.factor(dtest$Obesity_level),positive='1')
rf_cm
rf_cm$byClass
```
let's see what are the most important variables
```{r}
class_rf_impvars=varImpPlot(class_rf) 
```

