---
title: "Statistical learning- Unsupervised project"
author: "Mina Beric"
date: "2024-05-10"
output: html_document
---
Load the libraries
```{r setup, include=FALSE}
library(tidyverse)
library(pastecs)
library(readr)
library(ggplot2) 
library(factoextra) 
library(gridExtra) 
library(knitr)
library(dplyr) 
library(tidyr) 
library(caret)
library(reshape2)
library(corrplot)
library(gplots)
library(preprocessCore)
library(summarytools)
library(ggbiplot)
library(cluster)
```

Load the dataset 
```{r}
dataset<- read.csv("C:/Users/beric/Downloads/tarvel+review+ratings/google_review_ratings.csv")
glimpse(dataset)
```
Get rid of unecessary variables and rename the columsn to be more interpretable
```{r}

dataset<-dataset[,-1]
dataset <- dataset[, -ncol(dataset)]


new_names <- c('churches', 'resorts', 'beaches', 'parks', 'theatres',
               'museums', 'malls', 'zoo', 'restaurants','pubs_bars',
               'local_services','burger_pizza', 'hotels','juice_bars', 
               'art_galleries', 'dance_clubs', 'swimming_pools',
               'gyms', 'bakeries', 'beauty_spas', 'cafes', 'view_points',
               'monuments','gardens')
colnames(dataset) <- new_names
glimpse(dataset)
attach(dataset)
```
Change into numeric variable local_service

```{r}
dataset$local_services<-as.numeric(dataset$local_services)
```
```{r}
#Check for missing values

missing_values_count <- colSums(is.na(dataset))

total_missing_values <- sum(missing_values_count)

print(total_missing_values)
```
```{r}

dataset_clean <- na.omit(dataset)
anyNA(dataset_clean)

```
```{r}
#Check for duplicates

sum(duplicated(dataset_clean)) 
```
```{r}
#Get rid of them
dataset_clean<-distinct(dataset_clean)
tail(dataset_clean)
```
EXPLANATORY DATA ANALYSIS

```{r}
summary_stats <- summarytools::descr(dataset_clean, round.digits = 2, transpose = TRUE)
summary_stats
```
```{r}
summary(dataset_clean)
```
```{r}
# Reshape the data into long format
long_dataset <- tidyr::gather(dataset_clean, key, value)

# Plotting boxplots for each variable
ggplot(long_dataset, aes(x = key, y = value)) + 
  geom_boxplot(fill="skyblue", color="black", outlier.colour = "deepskyblue") + 
  facet_wrap(~key, scales = 'free')+
  theme_minimal()

#There are a lot of outliers, we need to handle this problem before clustering. 

#But first let's analyse more in detail our variables: 

ggplot(dataset_clean, aes(x = churches)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Churches", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = resorts)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Resorts", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = beaches)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Beaches", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = parks)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Parks", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = theatres)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Theaters", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = museums)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Museums", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = malls)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Malls", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = zoo)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Zoos", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = restaurants)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Restaurants", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())


ggplot(dataset_clean, aes(x = pubs_bars)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Pubs_bars", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = local_services)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Local_services", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = burger_pizza)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Burger pizza", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = hotels)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Hotels", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = juice_bars)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Juicy bars", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = art_galleries)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Art galleries", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = dance_clubs)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Dance clubs", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = swimming_pools)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Swimming pools", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = gyms)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Gyms", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = bakeries)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Bakeries", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = beauty_spas)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Beauty spas", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = cafes)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Cafes", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = view_points)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Viewv points", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = monuments)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Monuments", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid.minor = element_blank())

ggplot(dataset_clean, aes(x = gardens)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs( x = "Gardens", y = "Frequency")+
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey", linetype = "dashed"),
        panel.grid
        

```
Melt the dataframe to see the overall ratings distribution
```{r}
# Melt the dataframe
df_melted <- melt(dataset_clean)

# Create histogram
histplot_rating <- ggplot(data = df_melted, aes(x = value)) +
  geom_histogram(color = "black", fill = "lightgreen", bins = 30, alpha = 0.8) +
  labs(x = "Rating", y = "Count", title = "Ratings distribution") +
  theme_minimal() +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14, face = "bold"), plot.title = element_text(size = 18, face = "bold"))

# Display the plot
print(histplot_rating)
```
Then correlation between the variables
First let's see the heat map
```{r}
corr <- cor(dataset_clean)

mask <- matrix(FALSE, nrow = ncol(corr), ncol = ncol(corr))
mask[upper.tri(mask)] <- TRUE

#Heatmap
ggplot(data = melt(corr), aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "lightyellow", high = "darkblue") +
  labs(title = "Heatmap", x = "", y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank(),
        legend.position = "bottom")

corrplot(corr,type = "upper", method = "number", tl.col = "black",tl.srt = 45, tl.cex = 0.8, color = colorRampPalette(c("blue", "white", "red")))


#Cluster map
heatmap(corr, col = colorRampPalette(c("lightyellow", "darkblue"))(100), scale = "none", symm = TRUE)


```
Then Scaling the data: 
```{r}
dataset_scaled<-scale(dataset_clean)

```

PRINCIPAL COMPONENT ANALYSIS
```{r}
pcaCharts <- function(x) {
  x.var <- x$sdev ^ 2
  x.pvar <- x.var/sum(x.var)
  print("proportions of variance:")
  print(x.pvar)
  
  par(mfrow=c(2,2))
  plot(x.pvar,xlab="Principal component", ylab="Proportion of variance explained", ylim=c(0,1), type='b')
  plot(cumsum(x.pvar),xlab="Principal component", ylab="Cumulative Proportion of variance explained", ylim=c(0,1), type='b')
  screeplot(x)
  screeplot(x,type="l")
  par(mfrow=c(1,1))
}

# Perform PCA on the cleaned and imputed data
pca_result <- prcomp(dataset_scaled,center=TRUE, scale.=FALSE)
var <- get_pca_var(pca_result)
var$contrib   #Variables contributions 

kdf <- data.frame(pca_result$x[,1:2]) #First 2 dimensions 
kdf

# Print the summary of the PCA result
summary(pca_result)

pcaCharts(pca_result)

#biplot
biplot(pca_result, scale=0,cex=.7, )
pca_result$rotation=-pca_result$rotation
pca_result$x=-pca_result$x
biplot(pca_result, scale=0)

#variables
fviz_pca_var(pca_result)

#eigen values
eig<- get_eigenvalue(pca_result)

#scree plot explained variance
x11()
fviz_eig(pca_result, addlabels = TRUE)+theme_classic()

#cumulative explained variance
cs = cumsum(pca_result$sdev^2 / sum(pca_result$sdev^2))
cs
x11()
plot(cs, type="b", pch = 19,
     xlab = "# Dimensions", ylab = "% Variability", col = "black") 

abline(h = 0.95, col = "red", lty = 1)


text(10, 0.9, label = "95% cut-off threshold", col = "red", cex = 1.6)  
#let's secure 95% percent of explained variability 

n_components <- which.max(cs >= 0.95)
n_components #21 components 

dim1<-fviz_contrib(pca_result,choice="var",axes=1)
dim2<-fviz_contrib(pca_result,choice="var",axes=2)
x11()
grid.arrange(dim1, dim2, ncol = 2)


rho <- cor(dataset_scaled)
round(rho,3)

eigen(rho)

screeplot(princomp(dataset_scaled, cor=T))
```
K-MEANS CLUSTERING
```{r}
# Define cluster numbers to test
cl_list <- 1:10

# List to store inertia values
inertia_list <- vector(length = length(cl_list))

# Loop through cluster numbers
for (k in seq_along(cl_list)) {
  n_clusters <- cl_list[k]
  
  
  pca <- prcomp(dataset_scaled, center = TRUE, scale = FALSE)  
  data_reduced <- pca$x
  
  # KMeans clustering
  kmeans_model <- kmeans(data_reduced, centers = n_clusters, nstart = 10) 
  
  # Extract inertia (within-cluster sum of squares)
  inertia_list[k] <- sum(kmeans_model$withinss)
}

# Plot inertia vs number of clusters
plot(cl_list, inertia_list, type = "o", col = "black", xlab = "Number of Clusters", ylab = "Inertia")
#ELBOW AT THE 4TH CLUSTER 

# Set the number of clusters
k <- 4

# Run k-means clustering
kmeans_result <- kmeans(dataset_scaled, centers = k)

# Get cluster labels
clusterNames <- kmeans_result$cluster

# Perform PCA
pca_result <- prcomp(dataset_scaled, center = TRUE, scale. = FALSE)

# Get the first two principal components
X_r <- pca_result$x[, 1:2]

# Plot the clusters
plot(X_r, col = clusterNames, pch = 19,
     main = "Google dataset clusters using PCA")
legend("topright", legend = unique(clusterNames), col = 1:k, pch = 19, title = "Cluster")

```
```{r}
# Convert clusterNames to a list
people <- as.list(clusterNames)

# Initialize lists to store indices of points in each cluster
zero <- c()
one <- c()
two <- c()
three <- c()

# Loop through data and split points based on their clusters
for (i in seq_along(people)) {
  if (people[[i]] == 1) {
    zero <- c(zero, i)
  } else if (people[[i]] == 2) {
    one <- c(one, i)
  } else if (people[[i]] == 3) {
    two <- c(two, i)
  } else {
    three <- c(three, i)
  }
}


# Initialize a list to store the mean values for each cluster
mean_values <- list()

# Loop through each cluster
for (i in 1:k) {
  # Get the indices of points in the current cluster
  cluster_indices <- which(clusterNames == i)
  
  # Compute the mean of the rows corresponding to the current cluster
  mean_values[[i]] <- round(colMeans(dataset_scaled[cluster_indices, , drop = FALSE]), 4)
}

# Print the mean ratings for each cluster
for (i in 1:k) {
  cat("Cluster", i, ":", "\n")
  print(mean_values[[i]])
  cat("\n")
}

```

